---
title: "HW3"
author: "Tina"
date: "4/7/2018"
output: html_document
---

#######2.1 (a) DATA CLEAN UP (i) Read in txt as csv
```{r}
 rm(list=ls())
getwd()
setwd("/Users/tinapraveen/Desktop/HW3")
FILES <- list.files( pattern = ".txt")
for (i in 1:length(FILES)) {
FILE=read.table(file=FILES[i],header=T)
write.csv(FILE,file=paste0("/Users/tinapraveen/Desktop/HW3",sub(".txt","",FILES[i]),".csv"))
}
df <- read.csv("HW3file19.csv", header=TRUE , stringsAsFactors = FALSE)
df <- df[ ,-1]
```
#######2.1 (a)(ii) Does the data need to be standardized? 

YES

#######2.1 (a) (iii) Extra white spaces and delimiters are handled during conversion and the clean data set is uploaded.
```{r}
str(df)
names(df)
cols <- names(df)
cols.char <-"Name"
cols.num <- cols[!cols %in% cols.char]

df.char <- df[cols.char]
df.num <- as.data.frame(lapply(df[cols.num],as.numeric))
df.scaled <- scale(df.num)
df <- cbind(df.char, df.scaled)
str(df)
```
#######2.1 (b) CLUSTERING

```{r}
library(cluster)
library(factoextra)
#######2.1 (b) (i) Number of clusters is determined as 8 from the silhoutte graph
fviz_nbclust(df[2:9], kmeans, method="silhouette")
#######2.1 (b) (ii) Plot Clusters
fviz_cluster(kmeans(df[2:9], centers=8, nstart=25), data=df[2:9])
k <- kmeans(df[2:9], centers=8, nstart=25)
names(k)
print(k)
k$totss
k$withinss
k$size
```

#######2.1 (b) (iii)How many observations are in each cluster?
#K-means clustering with 8 clusters of sizes 9, 17, 2, 17, 9, 3, 1, 8
#######2.1 (b) (iv) What is the total SSE of the clusters?
# 520
#######2.1 (b)(v) What is the SSE of each cluster?
3.174224 14.918373  2.223560  9.591579  6.337449  4.546254  0.000000  4.244696

#######2.1 (b)(vi) Perform an analysis of each cluster to determine how the mammals are grouped in each cluster, and whether
that makes sense? 

The clusters do make sense .
for e.g: River Otter Sea Otter, Sea Lion etc clustered together and in another Cluster we have Porcupine,Pack Rat , Beaver etc being clustered and we have 8 clusters in total.

```{r}
for( i in 1:8){
  print(paste("Mammals in cluster", i))
cluster_mammals<- ((df[c(which(k$cluster == i)), 1] ))
print (cluster_mammals)
  
}

```
#############2.2 Problem 2: Hierarchical clustering

```{r}
getwd()
setwd("/Users/tinapraveen/Desktop/HW3")
FILES <- list.files( pattern = ".txt")
for (i in 1:length(FILES)) {
FILE=read.table(file=FILES[i],header=T)
write.csv(FILE,file=paste0("/Users/tinapraveen/Desktop/HW3",sub(".txt","",FILES[i]),".csv"))
}
df_46 <- read.csv("HW3file46.csv", header=TRUE, stringsAsFactors = FALSE)
str(df_46)
df_46 <- df_46[ ,-1]
row.names(df_46 )<-df_46$Country
df_46 <- scale(df_46[,2:13])
#df_46 <- df_46[ ,-1]
#str(df_46)
#names(df_46)
#cols <- names(df_46)
#cols.char <-"Country"
#cols.num <- cols[!cols %in% cols.char]
#df_46.char <- df_46[cols.char]
#df_46.num <- as.data.frame(lapply(df_46[cols.num],as.numeric))
#df_46.scaled <- scale(df_46.num)
#df_46 <- cbind(df_46.char, df_46.scaled)
#str(df_46)

#library(tidyverse)
#df_46 %>% remove_rownames %>% column_to_rownames(var="Country")

```
#############2.2 (a) Run Hierarchical Clustering
```{r}
#Single

hc.single <- eclust(df_46[ ,2:12] ,"hclust", hc_method="single")
fviz_dend(hc.single, show_labels= TRUE, palette="jco", as.ggplot=T)
# Average
hc.average <- eclust(df_46[ ,2:12] ,"hclust", hc_method="average")
fviz_dend(hc.average, show_labels= TRUE, palette="jco", as.ggplot=T)
# Complete
hc.complete <- eclust(df_46[ ,2:12] ,"hclust", hc_method="complete")
fviz_dend(hc.complete, show_labels= TRUE, palette="jco", as.ggplot=T)
clusters.5 <- cutree(hc.complete, 5 )
d<-dist(df_46[ ,2:12])
s <- cluster::silhouette(clusters.5, d)
summary(s)

```
#############2.2 (b)
Singleton Clusters in Linkage method: Single
Luxemburg and Switzerland speak GE in majority
West Germany and Austria speak GE  in majority
Great Britain and Ireland speak En in majority
This clustering makes sense.
Singleton Clusters in Linkage method: Average
Luxemburg and Switzerland speak GE in majority
West Germany and Austria speak GE  in majority
Great Britain and Ireland speak EN in majority 
Belgium and Finland speak EN in equal numbers 
This clustering makes sense.
Singleton Clusters in Linkage method: Complete
West Germany and Austria speak GE  in majority
Great Britain and Ireland speak EN in majority
Luxemburg and Switzerland speak GE in majority
Denmark and Norway speak EN in equal numbers though the language spoken by the majority in both these countries are different. 
Portugal and Spain speak EN in equal numbers though the language spoken by the majority in both these countries are different. 

#############2.2 (c)

#############2.2 (d)
Complete linkage method finds most number of singleton clusters and hence this wiould be considered pure.
#############2.2 (e)
Number of clusters after cutting at specified height is 5
#############2.2 (f)
```{r}
#Single
hc.single <- eclust(df_46[ ,2:12] ,k=5,"hclust", hc_method="single")
fviz_dend(hc.single, show_labels= TRUE, palette="jco", as.ggplot=T)
# Average
hc.average <- eclust(df_46[ ,2:12] ,k=5,"hclust", hc_method="average")
fviz_dend(hc.average, show_labels= TRUE, palette="jco", as.ggplot=T)
# Complete
hc.complete <- eclust(df_46[ ,2:12] ,k=5,"hclust", hc_method="complete")
fviz_dend(hc.complete, show_labels= TRUE, palette="jco", as.ggplot=T)

```
#############2.2 (g)
```{r}
library(fpc)
fpc::cluster.stats(d, hc.single$cluster)$avg.silwidth
fpc::cluster.stats(d, hc.average$cluster)$avg.silwidth
fpc::cluster.stats(d, hc.complete$cluster)$avg.silwidth
fpc::cluster.stats(d, hc.single$cluster)$dunn
fpc::cluster.stats(d, hc.average$cluster)$dunn
fpc::cluster.stats(d, hc.complete$cluster)$dunn


```
#############2.2 (h)
Considering the average silhouette width,  we will go for Complete linkage

#############2.2 (i)

Considering dunn index , we will go for Average linkage

##############2.3 Problem 3: K-Means and PCA
##############2.3 a 
```{r}
df_hrtu <- read.csv("HTRU_2-small.csv", header=TRUE, stringsAsFactors = FALSE)

pca <- prcomp(scale(df_hrtu))
pca
summary(pca)
##############2.3 a (ii)
biplot(pca, scale=0 ,col = c("blue","orange"))
raw <- pca$x[,1:2]
plot(raw[,1], raw[,2], col=rainbow(nrow(raw)), pch=20)

##############2.3 b
k <- kmeans(scale(df_hrtu[ ,1:8]), centers=2, nstart=25) 

fviz_cluster(k, data= scale(df_hrtu[ ,1:8]) )  
##############2.3 b  ii)
#The clusters are having the shape as the PCA plot. This cpuld occur because we have plotted the PCA agianst 2 dimensions which could explain maximum variance and hence the shapes are nearly same.
##############2.3 b  iii)) What is the distribution of the observations in each cluster?
k
# K-means clustering with 2 clusters of sizes 8847, 1153

#(iv) What is the distribution of the classes in the HTRU2 dataset?
library(plyr)
count(df_hrtu, "class")
#Class  Frequency
#0	    9041			
#1	    959	

#(v)
# Comparing above results cluster with size 8847 could belong to class 0
# and  cluster with size 1153 could belong to class 1

# (vi) Let’s focus on the larger cluster.
  cluster_obs <-which(k$cluster == 1)
  df_cluster1 <- df_hrtu[ cluster_obs , ]
  count(df_cluster1, "class")
  
  # Below is the frequency of classes in the larger cluster
  # class   freq 
#    0	    8624			
 #   1	    223
  #(vii) Based on the analysis above, which class (1 or 0) do you think the larger cluster represents?
  
#  Class 0
  
#  (viii) How much variance is explained by the clustering?
 # 35.9 %
  
#  (ix) What is the average Silhouette width of both the clusters?
dh<-  dist(scale(df_hrtu[ ,1:8]) )
fpc::cluster.stats(dh, k$cluster)$avg.silwidth
 # 0.6006794

#(x) What is the per cluster Silhouette width? Based on this, which cluster is good?
 sil_width <- silhouette(k$cluster , dh)
summary(sil_width)
?silhouette
#Cluster one is good.
```
##############2.3 a (i)
 cumulative variance is explained by the first two components -0.769
 
##############2.3 a (iii)
 
Plot forms 2 clusters based on class 1 and 0

##############2.3 C
```{r}
kpca <- kmeans(pca$x[, 1:2], centers=2, nstart=25) 
fviz_cluster(kpca, data= pca$x[ , 1:2])
###############2.3 C (i)
#All of the clusters have more or less same shape . This is because PCA could choose #the dimensions those could explain maximum variance within the dataset


#(ii) What is the average Silhouette width of both the clusters?
dpca<-  dist(pca$x[, 1:2])
fpc::cluster.stats(dpca, kpca$cluster)$avg.silwidth
#0.700875
#(iii) What is the per cluster Silhouette width? Based on this, which cluster is  good?
sil_width <- silhouette(kpca$cluster , dpca)
summary(sil_width)

# Cluster one is good.
#(iv) How do the values of c(ii) and c(iii) compare with those of b(ix) and b(x), respectively?

#c(ii) and c(iii)values are better
```
 