---
title: "HW1"
author: "Tina"
date: "2/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
2.1.A
```{r}
setwd( "/Users/tinapraveen/desktop")
college<-read.csv(file="college.csv",header=TRUE, sep=",")
```
2.1.B
```{r}

rownames(college)<- college[ ,1]
#fix(college)
college <- college [ , -1]
#fix(college)
```
When I try to use fix function in mac , it opens in XQuartz and then conversion to html is stopped.Hence the statement is commented.
2.1.C
```{r}
summary(college)
pairs(college[1:10])
boxplot(perc.alumni~ Private, data=college, main = "Alumni Donation" )
boxplot(PhD~ Private, data=college, main = "PhD Employ " )
Elite <- rep('No', nrow(college))
Elite[college$Top10perc > 50] <- 'Yes'
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)
summary(college)
```
iii)Alumni in Private schools donate more to their colleges.
iv)Public Schools employ more Ph.D.'s.
v)There are 78 Elite colleges.

```{r}
#range(college$Enroll)
#range(college$F.Undergrad)
#range(college$Personal)
#range(college$Expend)

par(mfrow=c(2,2))
hist(college$Enroll, 
     main="Histogram for Number of students Enrolled", 
     xlab="Applications", 
     border="blue", 
     col="green",
     breaks=seq(35,6500,100))
hist(college$F.Undergrad, 
     main="Histogram for Number of F.T Under grads", 
     xlab="FT UG", 
     border="blue", 
     col="green",
     breaks=seq(100,33000,500))
hist(college$Personal, 
     main="Histogram for Personal Spending", 
     xlab="Personal Spending", 
     border="blue", 
     col="green",
     breaks=seq(200,7000,100))
hist(college$Expend, 
     main="Histogram for     Expenditure", 
     xlab="Expenditure", 
     border="blue", 
     col="green",
     breaks=seq(3000,60000,1500))

 
```
2.2.a
```{r}
library(ggplot2)
library(GGally)
library(lattice)
nba<-read.csv(file="nba.csv",header=TRUE, sep=",")
str(nba)
ggcorr(nba, label = TRUE)
model1<-lm(PTS~FG, data = nba)
summary(model1)


xyplot(PTS ~ FG, data = nba,  
       panel = function(x, y) {
         panel.xyplot(x, y)
         panel.abline(lm(y ~ x))
       }, 
       xlab = "FG", 
       ylab = "PTS ",
       main = "Univariate linear regression")
```
We can see that FG and PTS are highly coorelated at  0.9 . Summary shows that R-squared is.9166 which means that the model can explain 91.6 % of the variance.
```{r}
set.seed(1122)
index <- sample(1:nrow(nba), 250)
train <- nba[index, ]
test <- nba[-index, ]
library(psych)
pairs.panels(train[18:23])
cor(nba$FG,nba$PTS)
cor(nba$X3P,nba$PTS)
cor(nba$FTA,nba$PTS)
cor(nba$TOT,nba$PTS)
cor(nba$X3PA,nba$PTS)
cor(nba$DR,nba$PTS)

model2 <- lm(PTS ~ FG + X3P + FTA + MIN , data=train)
summary(model2)
model3 <- lm(PTS ~ FG + X3P + FTA + X3PA + DR , data=train)
summary(model3)
model4 <- lm(PTS ~FG+FGA+X3P+FTA+TOT+A, data=train)
summary(model4)
plot(model4, 1) 
hist(model4$residuals, xlab = "Model Residuals", 
     main="NBA Residual Histogram")
x <- -7:7
lines(x, 240*dnorm(x, 0, sd(model4$residuals)), col=2)


```
In model 2, MIN is insignificant and R squared is 0.99.
In model 3 , X3PA and DR are insignificant and the next model, model4 uses the variables,FG,FGA,X3P,FTA,TOT,A and indicates that FGA , TOT and A are not significant.
R squared for the model is 0.9944 and high F-statistic indicate that the model fits the data well.


PREDICTION
```{r}
res <-predict(model4, test)
testdf <- data.frame(cbind(test$PTS , res))
testdf$equals <- NULL
testdf[testdf$V1 != testdf$res,"equals"] <- "1"
b <- sum(is.na(testdf$equals))
cat(paste(" Number of fitted values matched (exactly) the PTS in the test dataset:",b))
print(b)

n <- dim(test)[1]
p <- 6
RSS  <- sum((test$PTS - res)^2)
RSE  <- sqrt(1/(n-p-1)*RSS) # 
TSS  <- sum((test$PTS - mean(test$PTS))^2)
Fstat   <- ((TSS - RSS)/3)/(RSS/(n-p-1))
R.sq <- cor(test$PTS, res)^2
cat(paste("Manually calculated R-Sq. is", R.sq))
cat(paste(" RSE in the test dataset:",RSE))
cat(paste(" TSS in the test dataset:",TSS))
cat(paste(" F-statistic in the test dataset:",Fstat))
cat(paste(" RSS in the test dataset:",RSS))

```

 -
