---
title: "HW2"
author: "Tina"
date: "3/3/2018"
output: html_document
---


```{r}
library(rpart)
library(caret)
library(rpart.plot)
library(ROCR)

rm(list=ls())

getwd()


options("digits"=3)
############ Q:2.1
set.seed(1122)
# Load test and train dataset
adult_train <- read.csv("adult-train(1).csv", header=T, sep=",", comment.char = '#')
adult_test <- read.csv("adult-test(1).csv", header=T, sep=",", comment.char = '#')
head(adult_train)
```

Q.2.1.a

```{r}
adult_train_del <- which((adult_train$workclass=="?") | (adult_train$native_country=="?") | (adult_train$occupation=="?"))
adult_train <- adult_train[-(adult_train_del),]
adult_test_del <- which((adult_test$workclass=="?") | (adult_test$native_country=="?") | (adult_test$occupation=="?"))
adult_test <- adult_test[-(adult_test_del),]
```
QSTN 2.1.b
QSTN 2.1.b i-top three important predictors in the model? CAPITAL GAIN , OCCUPATION,EDUCATION
QSTN 2.1.b ii-The first split is done on which predictor?CAPITAL GAIN
What is the predicted class of the first node?<=50K 
What is the distribution of observations between the “<=50K” and “>50K” classes at this node? .75, .25

```{r}
model <- rpart(income ~ ., method="class", data=adult_train)
summary(model)
rpart.plot(model, extra=104, fallen.leaves = T, type=4)
```
QSTN 2.c
QSTN 2.c(i) What is the balanced accuracy of the model? 0.726 

(ii) What is the balanced error rate of the model?0.274

(iii) What is the sensitivity?0.948 Specificity?0.504
(iv) What is the AUC of the ROC curve-0.843

```{r}
pred <- predict(model, adult_test, type="class")
confusionMatrix(pred, adult_test[, 15])

pred.rocr <- predict(model, newdata=adult_test, type="prob")[,2]
f.pred <- prediction(pred.rocr, adult_test$income)
f.perf <- performance(f.pred, "tpr", "fpr")
plot(f.perf, colorize=T, lwd=3)
auc <- performance(f.pred, measure = "auc")
cat(paste("AUC: ", round(auc@y.values[[1]], 3)))
```
QSTN 2.D
```{r}
printcp(model)
rm.index <- sample(which(adult_train$income=="<=50K"), size=15145)
adult_train_us <- adult_train[-rm.index,]

table(adult_train$income)
table(adult_train_us$income)
```
QSTN 2.E
(i) In the training dataset, how many observations are in the class “<=50K”? How many are in the class “>50K”? 7508
QSTN 2.E.iii)
i) What is the balanced accuracy of this model? 0.803 
(ii) What is the balanced error rate of this model?0.197
(iii) What is the sensitivity?0.769 Specificity?0.838
(iv) What is the AUC of the ROC curve.0.845 Plot the ROC curve.

```{r}
modelus <- rpart(income ~ ., method="class", data=adult_train_us)
summary(modelus)

pred <- predict(modelus, adult_test, type="class")
confusionMatrix(pred, adult_test[, 15])

new.pred.rocr <- predict(modelus, newdata=adult_test, type="prob")[,2]
new.f.pred <- prediction(new.pred.rocr, adult_test$income)
new.f.perf <- performance(new.f.pred, "tpr", "fpr")
plot(new.f.perf, colorize=T, lwd=3)
new.auc <- performance(new.f.pred, measure = "auc")
cat(paste("AUC ", round(new.auc@y.values[[1]], 3)))

```
Question 2.2
(i) What is the balanced accuracy of the model?0.646
(ii) What is the accuracy of the model?0.824
(iii) What is the sensitivity  0.996 and 
specificity of the model?0.296
(iv) What is the response class distribution in the test dataset? (That is, how many observations are labeled “>50K” and how many are labeled “<=50K”)?11312 and  1096
(v) Given the response class distribution, does the sensitivity and specificity make sense?Yes
(vi)which is the most important variable and which is the least important one? CAPITAL GAIN and NATIVE COUNTRY

```{r}
library(randomForest)
set.seed(1122)

rf.model <- randomForest(income ~ ., data=adult_train, importance=T)
rf.pred <- predict(rf.model, adult_test, type="class")
confusionMatrix(rf.pred, as.factor(adult_test$income))

varImpPlot(rf.model)

print(rf.model)

mtry <- tuneRF(adult_train[,-15], adult_train[,15], ntreeTry=500, stepFactor=1.5,improve=0.01, 
              trace=TRUE, plot=TRUE)
print(mtry)

rf.model_tune <- randomForest(income ~ ., data=adult_train, importance=T, mtry=2)
rf.pred_tune <- predict(rf.model_tune, adult_test, type="class")
confusionMatrix(rf.pred_tune, adult_test[,15])
```
Question 2.3
```{r}

library(arules)
transdata <- read.transactions("groceries.csv",sep=",")
summary(transdata)
#(i)
rules <- apriori(transdata)
summary(rules)
inspect(rules)
#(ii)At support count 0.001 we get atleast 400 rules
rules <- apriori(transdata, parameter = list(support=0.001))
summary(rules)
inspect(rules)
#(iii) Which item is the most frequently bought and what is its frequency?

# whole milk is the frequently brought item 
par(mfrow=c(1,2))

itemFrequencyPlot(transdata,
                  type="relative",
                  topN=10, # can be changed to the number of interest
                  horiz=TRUE,
                  col='steelblue3',
                  xlab='',
                  main='Item frequency, relative')

#(iv) Which item is the least frequently bought and what is its frequency?
#baby food and sound storage system both are infrequent.

par(mar=c(2,10,2,2), mfrow=c(1,2))

barplot(sort(table(unlist(LIST(transdata))))[1:10]/9835,
        horiz=TRUE,
        las=1,
        col='steelblue3',
        xlab='',
        main='Frequency, relative')

#(v) What are the top 5 rules, sorted by support?
inspect(sort(rules, by='support', decreasing = T)[1:5])

#(vi) What are the top 5 rules, sorted by confidence?
inspect(sort(rules, by='confidence', decreasing = T)[1:5])

#(vii) What are the bottom 5 rules, sorted by support?
inspect(sort(rules, by='support', decreasing = F)[1:5])

#(viii) What are the bottom 5 rules, sorted by confidence?
inspect(sort(rules, by='confidence', decreasing = F)[1:5])


````

